---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# herpesdrg

<!-- badges: start -->
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.4.0-6666ff.svg)](https://cran.r-project.org/)
![GitHub repo size](https://img.shields.io/github/repo-size/ojcharles/herpesdrg.svg)
![GitHub code size in bytes](https://img.shields.io/github/languages/code-size/ojcharles/herpesdrg.svg)
<!-- badges: end -->



## Overview

herpesdrg is a R package to enable antiviral drug resistance genotyping, with Herpes Simplex Virus 1, HSV 2, Human Cytomegalovirus and Varicella zoster virus sequencing data. Accepted inputs are FASTA (whole genomes & fragments) which will be mapped to RefSeq NC_001806.2. NGS variant data assembled to NC_001806.2 is accepted in VCF >= ver4.0 & Varscan2 tab formats.


#### Database
All data extracted from literature sources and contains the relationships between:

 - Virus
 - Gene
 - Mutation
 - Drug susceptibility fold change from sensitive strain or "Resistant" / "Polymorphism" as in publication
 - The method for Resistance Phenotyping and strain generation
 - Publication reference 
 

#### Web service
A user-friendly Shiny Applications has been bundled with this package. The same application is available over the internet here [http://cmv-resistance.ucl.ac.uk/herpesdrg/](http://cmv-resistance.ucl.ac.uk/herpesdrg/) where the terms of use are contained.


## Installation

You can install the current version from [GitHub](https://github.com/ojcharles/herpesdrg) with:

``` r
# install.packages("devtools")
devtools::install_github("ojcharles/herpesdrg")
```

Dependencies for FASTA file handling are MAFFT and SNP-Sites available preferably via conda. snp-sites >= 2.3 has been tested.
```{bash, eval=FALSE}
conda config --add channels bioconda
conda install snp-sites
conda install mafft
```


## Usage

### vcf data

```{r setup, message = F}
library("herpesdrg")

## call resistant variants
my_sample = system.file("testdata", "HSV1_F716L.vcf", package = "herpesdrg")

data = call_resistance(infile = my_sample, all_mutations = T, virus = "HSV1") # options are HSV1 HSV2 HCMV and VZV

data[ , c("change", "freq", "Aciclovir", "Pencyclovir", "Foscarnet")]


## call all variants
mutations_all = call_resistance(infile = my_sample, all_mutations = T)

#to view all mutations in resistance genes we can filter
mutations_res = mutations_all[mutations_all$GENEID %in% c("UL30"),]

head(mutations_res[,c(1,8,21,32:40)])


## call nonsynonymous variants
# are there any non-synonymous (DNA variants that result in a change of amino acid) variants in resistance genes
mutations_res_nonsyn = mutations_res[mutations_res$CONSEQUENCE == "nonsynonymous",]


# here the top 3 mutations are nonsynonymous, with no identified resistance effect.
head(mutations_res_nonsyn[,c(1,8,21,32:40)])


```


### fasta sequences
```{r}
# herpesdrg accepts sequence fragments or whole-genomes, one fasta sequence at a time.

# load example data
my_sequence = system.file("testdata", "HSV1_F716L.vcf", package = "herpesdrg")

dat = call_resistance(infile = my_sequence, all_mutations = T)

head(dat)
```


### other features
```{r}
## view the full database
db = herpesdrg_data()
head(db$aa_change)


## run the shiny application
# ShinyHerpes()

```

## Getting help

If you encounter a clear bug, please file an issue with a minimal reproducible example on the GitHub Issues page. For questions and other discussions feel free to contact. [Oscar Charles - maintainer](mailto:oscar.charles.18@ucl.ac.uk)

